### 2018-05-15 1_intro.tex
consider some Professor's comments.

\begin{figure}[t]
\centering
\ttfamily
\begin{tabular}{ |l|l| }
\hline
\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P0 & P1 \\ \hline
MOV [x],1 & MOV [y],1 \\
MOV EAX,[y] & MOV EBX,[x] \\
\hline
%\multicolumn{2}{|l|}{locations [x;y;]} \tabularnewline
\multicolumn{2}{|l|}{exists (0:EAX=0,~1:EBX=0)} \tabularnewline
\hline
\multicolumn{2}{|l|}{x86: allow} \tabularnewline
\hline
\end{tabular}
\caption{Store buffering: A litmus test on write-read reordering allowed under the x86-TSO and forbidden under the SC memory model}
\label{simple_wmm_x86}
\end{figure}

@@

In addition to developing the theoretical basis, researchers work on extracting the memory models for hardware architectures from existing implementations or from the specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
Over the last decade the memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc. %, or Alpha architecture~\cite{alphaWMM}~, and others.
There are projects for validating hardware architectures wrt. a memory model as well, e.g.~\cite{lustig2014pipecheck,lustig2016coatcheck}.

@@

All the tools listed above consider only a single memory model, however, in real life we face serious engineering problems involving necessity to model more than one execution environment.
One of these problems is the \textit{portability} of the program from one hardware architecture to another.
A program written in a high-level language is then compiled for different hardware.
Even if all the compiler optimisations were disabled (which is rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
As the result, a program compiled under the platforms $\mathcal{T}$ can reach states that are unreachable on the platform $\mathcal{S}$, which is a \textit{portability bug} from the source platform $\mathcal{S}$ to the target platform $\mathcal{T}$~\cite{Porthos17a}.

The very first tool that performs the WMM-aware portability analysis is \porthos%
%
\footnote{The \porthos{} project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}} %
%
introduced in April~2017~\cite{Porthos17b}.
This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT-solver.
This approach allows to capture symbolically the semantics of analysing program and both weak memory models into a single SMT-formula, augmented by the reachability assertion.
As most modern SMT-solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
% TODO: search space size: https://courses.cs.washington.edu/courses/csep573/11wi/lectures/ashish-satsolvers.pdf slide

@@

The current work aims to rework the proof-of-concept tool \textit{\porthos{}} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
One of the directions of development was the ability to process the \textit{kernel litmus tests} written in C~\cite{TODO: KERNEL LITMUS TESTS WORK+REPOSITORY}.
%interprocedural analysis
As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name \textit{\porthos[2]}\,%
%
\footnote{Hereinafter with the name `\porthos{}' we refer to the tool \porthos{} of version 1, whereas the new implementation of \porthos{} is called \porthos[2].}. %
%
Considering the enhancements of the architectural design, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static code analysis.

@@

The thesis is organised as following.
Chapter~\ref{ch:wmm} gives a general view on the weak memory model-aware analysis.
Chapter~\ref{ch:port} examines the portability analysis as an bounded reachability problem that can be encoded into an SMT-formula in order to be solved automatically by an SMT~solver.
Chapter~\ref{ch:impl} delves into the description of architectural solutions and implementation details of the \porthos[2] framework.
\textbf{TODO: other chapters!!!}

%Clarke's goto-cc : \cite{clarke2004tool} : http://www.cprover.org/goto-cc/
% (perhaps, goto-cc is a part of cprover) : cprover https://www.cprover.org/cbmc/doc/manual.pdf
%cbmc http://www.cprover.org/cbmc/

>>>>

\begin{figure}[t]
\centering
\ttfamily
\begin{tabular}{ |l|l| }
\hline
\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P0 & P1 \\ \hline
MOV [x],1 & MOV [y],1 \\
MOV EAX,[y] & MOV EBX,[x] \\
\hline
%\multicolumn{2}{|l|}{locations [x;y;]} \tabularnewline
\multicolumn{2}{|l|}{exists (0:EAX=0,~1:EBX=0)} \tabularnewline
\hline
\multicolumn{2}{|l|}{x86: allow} \tabularnewline
\hline
\end{tabular}
\caption{Store buffering: A litmus test illustrating the write-read reordering allowed by the x86-TSO memory model}
\label{simple_wmm_x86}
\end{figure}

@@

In addition to developing the theoretical basis, researchers work on extracting the memory models for hardware architectures from existing implementations or from the specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
Over the last decade, memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc. %, or Alpha architecture~\cite{alphaWMM}~, and others.
There are projects for validating hardware architectures wrt. a memory model as well, e.g.~\cite{lustig2014pipecheck,lustig2016coatcheck}.

@@

All the tools listed above consider only a single memory model, however, in real life we face serious engineering problems involving necessity to model more than one execution environment.
One of these problems is the \textit{portability} of the program from one hardware architecture to another.
A program written in a high-level language is then compiled for different hardware.
Even if all the compiler optimisations were disabled (which is rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
As the result, a program compiled under the platform $\mathcal{T}$ can reach states that are unreachable on the platform $\mathcal{S}$, which is a \textit{portability bug} from the source platform $\mathcal{S}$ to the target platform $\mathcal{T}$~\cite{Porthos17a}.

The very first tool that performs the WMM-aware portability analysis is \porthos%
%
\footnote{The \porthos{} project web site: \url{http://github.com/hernanponcedeleon/Dat3M}} %
%
introduced in April~2017~\cite{Porthos17b}.
This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT-solver.
This approach allows to capture symbolically the semantics of analysing program and both weak memory models into a single SMT-formula, augmented by the reachability assertion.
As most modern SMT-solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
% TODO: search space size: https://courses.cs.washington.edu/courses/csep573/11wi/lectures/ashish-satsolvers.pdf slide

@@

The current work aims to rework the proof-of-concept tool \textit{\porthos{}} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
One of the directions of development was the ability to process the \textit{kernel litmus tests} written in C~\cite{mckenney2017wg21}.
For that, in addition to supporting the C syntax (augmented by litmus-style definitions such as initialisation or assertion statements), the tool must be able to recognise kernel-specific functions and macros (such as the macro \texttt{READ\_ONCE} that guarantees memory read).
%interprocedural analysis

As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name \textit{\porthos[2]}\,%
%
\footnote{Hereinafter with the name `\porthos{}' we refer to the tool \porthos{} of version 1, whereas the new implementation of \porthos{} is called \porthos[2].}. %
%
Considering the enhancements of the architectural design, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static analysis of concurrent programs.

@@

The thesis is organised as following.
Chapter~\ref{ch:wmm} gives a general view on the weak memory model-aware analysis.
Chapter~\ref{ch:port} examines the portability analysis as an bounded reachability problem that can be encoded into an SMT-formula in order to be solved automatically by an SMT~solver.
Chapter~\ref{ch:impl} delves into the description of architectural solutions and implementation details of the \porthos[2] framework.
Chapter~\ref{ch:eval} provides an example of the portability analysis via \porthos[2] and presents some evaluation results comparing to other tools. %TODO: check if it's true
Chapter~\ref{ch:summary} summarises results of the work and proposes future work directions. %TODO: check if it's true

### 2018-05-15 2_wmm.tex
consider some Professor's comments.

The main idea behind the memory model-aware program analysis is that the set of all possible interleavings of the concurrent program (\textit{anarchic semantics}) can be specified by the axiomatic constraints of the memory model that filter out executions inconsistent in particular architecture (\textit{analytic semantics})~\cite{alglave2016syntax}.
The anarchic semantics of the program is a truly parallel semantics with no global time that describes all possible computations with all possible communications.
However, the analytic semantics captures the program behaviours on a certain execution environment more precisely.
%"The semantics of a program is a set of executions" https://johnwickerson.github.io/papers/memalloy.pdf

@@

There are three main types of sources of non-determinism in concurrent programs~\cite{musuvathi2008fair}:
\begin{enumerate}[noitemsep,topsep=0pt]
\item \textit{input non-determinism}, which is a standard undecidable problem for all static analysis methods: to resolve the user input, system call from the environment, unresolved function calls, etc.;
\item \textit{scheduling non-determinism}, caused by the interleavings, which in turn are caused by the scheduler activity;
\item \textit{memory-model non-determinism}, caused by hardware and compiler relaxations.
\end{enumerate}

@@

    \item The \textit{data-flow} of a program is defined by \textit{communication relations}:
    \begin{itemize}[noitemsep]
        \item the \textit{read-from} relation \rf{}~$\subseteq~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event;
        \item the \textit{coherence order} relation \co{}~$\subseteq~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation);
    \end{itemize}

@@

    \item the \textit{from-read} relation $\fr \subseteq~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to all write events succeeding the write event from which the read event gets its value: \\
    $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$

@@

    \item the \textit{communication} relation \com over memory events, that fully describes the data-flow of a program: \\
    %$\mathtt{com} \triangleq (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
    $m_1 \relation{com} m_2~\triangleq~((m_1\relation{rf}m_2) \lor (m_1\relation{co}m_2) \lor (m_1\relation{fr}m_2))$

    \item the \textit{external} (and \textit{internal}) \textit{from-read} relations that restrict the \fr-relation to the different (respectively, same) processes: \\
    $w \relation{fre} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) = \mathtt{proc}(w))$ \\
    $w \relation{fri} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) \neq \mathtt{proc}(w))$

    %TODO: say that po-relation is *static*
    \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable: \\
    $m_1 \relation{po-loc} m_2 ~\triangleq~ (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$

@@

\begin{figure}[!htb]
\centering
\begin{subfigure}[t]{.28\textwidth}
  \centering
  \includegraphics[width=1.2\linewidth]{img/my/sb-example/SB-dot.png}
  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=1)}}
  \label{simple_wmm_x86_pic:sub1}
\end{subfigure}
\hfill
\begin{subfigure}[t]{.23\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-2.png}
  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=0)}}
  \label{simple_wmm_x86_pic:sub2}
\end{subfigure}
\hfill
\begin{subfigure}[t]{.23\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-3.png}
  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=1)}}
  \label{simple_wmm_x86_pic:sub3}
\end{subfigure}
\hfill
\begin{subfigure}[t]{.23\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-4.png}
  \caption{Final state: \texttt{(0:EAX=0,~1:EAX=0)}}
  \label{simple_wmm_x86_pic:sub4}
\end{subfigure}
\hfill
\caption{Possible candidate executions for the litmus test Example~\ref{simple_wmm_x86}}
\label{simple_wmm_x86_pic}
\end{figure}

@@

\section{\cat{} language}
\label{ch:wmm:cat}

>>>>

The main idea behind the memory model-aware program analysis is that the set of all possible interleavings of the concurrent program (the \textit{anarchic semantics}) can be specified by the axiomatic constraints of the memory model that filter out executions inconsistent in particular architecture (the \textit{analytic semantics})~\cite{alglave2016syntax}.
The anarchic semantics of the program is a truly parallel semantics with no global time that describes all possible computations with all possible communications.
However, the analytic semantics captures the program behaviours on a certain execution environment more precisely.
%"The semantics of a program is a set of executions" https://johnwickerson.github.io/papers/memalloy.pdf

@@

There are three main types of sources of non-determinism in concurrent programs~\cite{musuvathi2008fair}:
\begin{enumerate}[noitemsep,topsep=0pt]
\item \textit{input non-determinism}, which is a standard undecidable problem for all static analysis methods: to resolve the user input, system call from the environment, unresolved function calls, etc.;
\item \textit{scheduling non-determinism}, caused by the interleavings, which in turn are caused by the scheduler activity; and
\item \textit{memory-model non-determinism}, caused by hardware and compiler relaxations.
\end{enumerate}

@@

    \item The \textit{data-flow} of a program is defined by \textit{communication relations}:
    \begin{itemize}[noitemsep]
        \item the \textit{read-from} relation \rf{}~$\subseteq~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event; and
        \item the \textit{coherence order} relation \co{}~$\subseteq~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation).
    \end{itemize}

@@

    \item the \textit{from-read} relation $\fr \subseteq~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to all write events succeeding the write event from which the read event gets its value: \\
    $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$;

@@

    \item the \textit{communication} relation \com over memory events, that fully describes the data-flow of a program: \\
    %$\mathtt{com} \triangleq (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
    $m_1 \relation{com} m_2~\triangleq~((m_1\relation{rf}m_2) \lor (m_1\relation{co}m_2) \lor (m_1\relation{fr}m_2))$;

    \item the \textit{external} (and \textit{internal}) \textit{from-read} relations that restrict the \fr-relation to the different (respectively, same) processes: \\
    $w \relation{fre} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) = \mathtt{proc}(r))$, \\
    $w \relation{fri} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) \neq \mathtt{proc}(r))$;

    %TODO: say that po-relation is *static*
    \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable: \\
    $m_1 \relation{po-loc} m_2 ~\triangleq~ (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$; and

@@

\begin{figure}[!htb]
\centering
\begin{subfigure}[t]{.28\textwidth}
  \centering
  \includegraphics[width=1.2\linewidth]{img/my/sb-example/SB-dot.png}
  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=1)}}
  \label{simple_wmm_x86_pic:sub1}
\end{subfigure}
\hfill
\begin{subfigure}[t]{.23\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-2.png}
  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=0)}}
  \label{simple_wmm_x86_pic:sub2}
\end{subfigure}
\hfill
\begin{subfigure}[t]{.23\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-3.png}
  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=1)}}
  \label{simple_wmm_x86_pic:sub3}
\end{subfigure}
\hfill
\begin{subfigure}[t]{.23\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-4.png}
  \caption{Final state: \texttt{(0:EAX=0,~1:EAX=0)}}
  \label{simple_wmm_x86_pic:sub4}
\end{subfigure}
\hfill
\caption{Candidate executions for the litmus test in Example~\ref{simple_wmm_x86}}
\label{simple_wmm_x86_pic}
\end{figure}

@@

\section{The \cat{} language}
\label{ch:wmm:cat}

### 2018-05-15 3_encod.tex
write up to Interpreter.
consider some Professor's comments.

While working on the \porthos[2], we applied some modifications of the encoding scheme for the control-flow.
The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of C language.
For that, we compile the recursive abstract syntax tree~(AST) of the parsed C-code to the plain (non-recursive) event-flow graph.
We show %TODO! to show, actually!!!
that the new encoding is smaller than the old one used in \porthos{} since it does not produces new variables for each high-level statement of the input language.
%The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner. This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions. 
For instance, \porthos{} uses the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
$\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} = (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
and control-flow of the branching instruction $i_1 = (c \; ? \; i_2 \; : \; i_3)$ was encoded as
$\phi_{CF}( \; c \; ? \; i_2 \; : \; i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\mathtt{if}\;x\;\mathtt{then}\;y\;\mathtt{else}\;z$').
In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula. The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:port:enc:df}).

@@

In an SMT-formula, the relation $x \relation{r} y$ is represented by a boolean variable $r(x, y)$ that indicates whether the relation holds.
The derived relations are encoded by fresh boolean variables according to the following rules~\cite{Porthos17b}:
\begin{itemize}[noitemsep,topsep=0pt]
\item $r_1 \cup r_2(e_1, e_2) \defeq r_1(e_1, e_2) \lor r_2(e_1, e_2)$
\item $r_1 \cap r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land r_2(e_1, e_2)$
\item $r_1 \relminus r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land \neg r_2(e_1, e_2)$
\item $r^{-1}(e_1, e_2) \defeq r(e_2, e_1)$
\item $r^{*}(e_1, e_2) \defeq r^{+}(e_1, e_2) \lor (e_1 = e_2)$
\item $r_1;r_2(e_1,e_2) \defeq \bigvee\limits_{e_k \in \mathbb(E)}^{} r_1(e_1,e_k) \land r_2(e_k,e_2)$
\item $r^{+}(e_1, e_2) \defeq \mathtt{tc}_{\left\lceil log|\mathbb{E}| \right\rceil} (e_1, e_2), where \\
\hspace*{1em} \mathtt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), and \\ 
\hspace*{1em} \mathtt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\mathtt{tc}_{i}(e_1,e_2);\mathtt{tc}_{i}(e_1,e_2))$
\end{itemize}

Note that the \cat{} language allows mutually-recursive definitions of relations (for example, `$\mathtt{r_1} = \mathtt{r_2} \cup (r_1;r_1)$').
The basic idea of using the Kleene fixpoint iteration for encoding such relations was also proposed in~\cite{Porthos17a}: for any pair of events $e_1,e_2 \in \mathbb{E}$ and relation $r \subseteq \mathbb{E} \times \mathbb{E}$, we encode a new integer variable $\Phi_{e_1,e_2}^{r}$ that represents the round of Kleene iteration on which the variable $r(e_1, e_2)$ has been set.
 %is used for guessing the execution

>>>>

While working on the \porthos[2], we have applied some modifications to the encoding scheme for the control-flow.
The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
For that, we compile the abstract syntax tree~(AST) of the parsed C-code to the plain event-flow graph.
We show %TODO! to show, actually!!!
that the new encoding is smaller than the old one used in \porthos{} since it does not produces new variables for each high-level statement of the input language.
%The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner. This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions. 
For instance, \porthos{} used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
$\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} = (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
and control-flow of the branching instruction $i_1 = (c \; ? \; i_2 \; : \; i_3)$ was encoded as
$\phi_{CF}( \; c \; ? \; i_2 \; : \; i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\mathtt{if}\;x\;\mathtt{then}\;y\;\mathtt{else}\;z$').
In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula. The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:port:enc:df}).

@@

In an SMT-formula, the relation $x \relation{r} y$ is represented by a boolean variable $r(x, y)$ that indicates whether the relation holds.
The derived relations are encoded by fresh boolean variables according to the following rules~\cite{Porthos17b}:
\begin{itemize}[noitemsep,topsep=0pt]
\item $r_1 \cup r_2(e_1, e_2) \defeq r_1(e_1, e_2) \lor r_2(e_1, e_2)$;
\item $r_1 \cap r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land r_2(e_1, e_2)$;
\item $r_1 \relminus r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land \neg r_2(e_1, e_2)$;
\item $r^{-1}(e_1, e_2) \defeq r(e_2, e_1)$;
\item $r^{*}(e_1, e_2) \defeq r^{+}(e_1, e_2) \lor (e_1 = e_2)$;
\item $r_1;r_2(e_1,e_2) \defeq \bigvee\limits_{e_k \in \mathbb{E}}^{} r_1(e_1,e_k) \land r_2(e_k,e_2)$; \ and
\item $r^{+}(e_1, e_2) \defeq \mathtt{tc}_{\left\lceil log|\mathbb{E}| \right\rceil} (e_1, e_2), \ where \\
\hspace*{1em} \mathtt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), \ and \\ 
\hspace*{1em} \mathtt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\mathtt{tc}_{i}(e_1,e_2);\mathtt{tc}_{i}(e_1,e_2))$.
\end{itemize}

Note that \cat{} language allows mutually-recursive definitions of relations (for example, `$\mathtt{r_1} = \mathtt{r_2} \cup (r_1;r_1)$').
The basic idea of using the Kleene fixpoint iteration for encoding such relations was also proposed in~\cite{Porthos17a}: for any pair of events $e_1,e_2 \in \mathbb{E}$ and relation $r \subseteq \mathbb{E} \times \mathbb{E}$, we encode a new integer variable $\Phi_{e_1,e_2}^{r}$ that represents the round of Kleene iteration on which the variable $r(e_1, e_2)$ has been set.
 %is used for guessing the execution

### 2018-05-15 4_impl.tex
consider some Professor's comments.

The main call for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
This implies the support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
Yet far from all of C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
%
\footnote{For ensuring this, one merely has to look at existing C compilers, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
%
), we consider the accomplished work as a step towards it.
%This applies not only to instances that simply constitute the syntactic sugar of C language (such as the \texttt{switch} statement or the prefix increment operator \texttt{++i}), but also to its fundamental concepts and features (such as pointer arithmetic or structures).
%However, the architecture of current version of \porthos{} does not allow to ... as it is ....
% Although and optimisation of the tool so that it is able to process real C programs in perspective.

@@

\vspace{0.5em}
\begin{enumerate}[nolistsep]
  \item \textit{Robustness:}
    \begin{enumerate}[label*=\arabic*.]
      \item prohibit under-approximating abstractions;
      \item preservation the completeness of analysis;
      \item modular architecture: each module can be tested independently;
      \item usage of software design patterns where necessary;
      \item usage of immutable data structures for all DTOs;
    \end{enumerate}
  \item \textit{Transparency:}
    \begin{enumerate}[label*=\arabic*.]
      \item following the principles of simplicity and readability;
      \item clear and informative program output;
    \end{enumerate}
  \item \textit{Efficiency:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
      \item keeping the trade-off between execution time and memory usage;
    \end{enumerate}
  \item \textit{Extensibility:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1em]
      \item clear modular architecture.
    \end{enumerate}
\end{enumerate}

@@

>>>>

The main call for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
Yet far from all of C language are supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
%
\footnote{For ensuring this, one merely has to look at existing C compilers, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
%
), we consider the accomplished work as a step towards it.
%This applies not only to instances that simply constitute the syntactic sugar of C language (such as the \texttt{switch} statement or the prefix increment operator \texttt{++i}), but also to its fundamental concepts and features (such as pointer arithmetic or structures).
%However, the architecture of current version of \porthos{} does not allow to ... as it is ....
% Although and optimisation of the tool so that it is able to process real C programs in perspective.

@@

\vspace{0.5em}
\begin{enumerate}[nolistsep]
  \item \textit{Robustness:}
    \begin{enumerate}[label*=\arabic*.]
      \item preservation the completeness of analysis,
      \item modular architecture: each module can be tested independently,
      \item usage of software design patterns where necessary, and
      \item usage of immutable data structures for all DTOs;
    \end{enumerate}
  \item \textit{Transparency:}
    \begin{enumerate}[label*=\arabic*.]
      \item following the principles of simplicity and readability, and
      \item clear and informative program output;
    \end{enumerate}
  \item \textit{Efficiency:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
      \item keeping the trade-off between execution time and memory usage;
    \end{enumerate}
  \item \textit{Extensibility:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1em]
      \item clear modular architecture.
    \end{enumerate}
\end{enumerate}

@@

%TODO: say about IDs from old code
